---
output: rmarkdown::github_document
---

# Extracting soft skills and comparing resume frequency between cities
## Heather Geiger

## Load libraries and previous Rdata file.

```{r, echo=TRUE,eval=FALSE}
library(stringr)    #For string operations
library(rvest)      #For screen scrapper
library(tidyverse)  #For Tidyverse
library(RCurl)      #For File Operations
library(dplyr)      #For Manipulating the data frames
library(DT)         #For Data table package
library(curl)

library(RJSONIO) #For processing JSON format data

library(ggplot2) #For plotting
```

```{r, echo=FALSE,eval=TRUE}
library(stringr)
suppressWarnings(suppressMessages(library(rvest)))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(RCurl)))
library(dplyr)
suppressWarnings(suppressMessages(library(DT)))
suppressWarnings(suppressMessages(library(curl)))
library(RJSONIO)
```

```{r, echo=TRUE,eval=TRUE}
load("resumes_across_cities_one_text_field_per_resume.Rdata")
```

## Using regular expressions to count resumes containing each skill

Set up a dictionary of keywords per skill.

```{r, echo=TRUE,eval=TRUE}
keywords <- data.frame(Skill = c("Collaboration",
				"Communication",
				"Creativity",
				"Customer service",
				"Decision making",
				"Leadership",
				"Presentation",
				"Problem solving"),
			Synonyms = c("collaboration,collaborative,collaborate,collaborated,team player,teamwork",
				"communication,communicative,communicate,communicated",
				"creativity,creative",
				"customer service",
				"decision making,made decisions,making decisions,decision maker",
				"leadership,led,leader,leading",
				"presentation,presenting,presented",
				"problem solving,solving problems,solved problems"),
			stringsAsFactors=FALSE)

keyword_list <- vector("list",length=nrow(keywords))

for(i in 1:nrow(keywords))
{
keyword_list[[i]] <- unlist(strsplit(keywords$Synonyms[i],",")[[1]])
}

keyword_list
```

Write a function to give a pattern for a keyword if it has a word boundary, comma, or space on each side.

Then if there are multiple keywords, paste these together with a pipe.

Finally, run this function for every item in keyword_list.

```{r}
#Couldn't figure out how to get a regex for a space, comma, or word boundary. However did get one that can get either a space or comma.
space_or_comma <- "[[:space:],]"
word_boundary <- "\\b"

pattern_for_one_keyword <- function(keyword){
        regexes <- paste0(space_or_comma,keyword,space_or_comma)
        regexes <- c(regexes,paste0(word_boundary,keyword,word_boundary))
        regexes <- c(regexes,paste0(word_boundary,keyword,space_or_comma))
        regexes <- c(regexes,paste0(space_or_comma,keyword,word_boundary))
        return(paste0(regexes,collapse="|"))
}

pattern_for_multiple_keywords <- function(keyword_vector){
        if(length(keyword_vector) == 1){return(pattern_for_one_keyword(keyword_vector))}
        if(length(keyword_vector) > 1){
                individual_regexes <- c()
                for(i in 1:length(keyword_vector))
                {
                        individual_regexes <- c(individual_regexes,pattern_for_one_keyword(keyword_vector[i]))
                }
        return(paste0(individual_regexes,collapse="|")) 
        }
}

keyword_regexes <- unlist(lapply(keyword_list,function(x)pattern_for_multiple_keywords(x)))
```

We can now use keyword_regexes along with str_detect to give a TRUE/FALSE value for whether the text contains the pattern.

```{r}
num_resumes_per_skill <- c()

for(i in 1:length(keyword_regexes))
{
skill <- keyword_regexes[i]
skill_in_text <- str_detect(tolower(resumes_across_cities$Text),skill)
resumes_across_cities_incl_this_skill <- resumes_across_cities[skill_in_text,]
num_resumes_this_skill <- data.frame(resumes_across_cities_incl_this_skill %>% group_by(City) %>% summarize(Num.resumes = n()),stringsAsFactors=FALSE)
num_resumes_this_skill <- data.frame(Skill = keywords$Skill[i],num_resumes_this_skill,stringsAsFactors=FALSE)
if(i == 1){num_resumes_per_skill <- num_resumes_this_skill}
if(i > 1){num_resumes_per_skill <- rbind(num_resumes_per_skill,num_resumes_this_skill)}
}

num_resumes_per_skill
```

Finally, add a column Percent.resumes which expresses the percentage of resumes from the city that contain the skill.

```{r}
resumes_per_city <- data.frame(table(resumes_across_cities$City))
resumes_per_city[,1] <- as.vector(resumes_per_city[,1])

colnames(resumes_per_city) <- c("City","Total.resumes.this.city")

num_resumes_per_skill <- merge(num_resumes_per_skill,resumes_per_city,"City") %>% mutate(Percent.of.resumes.this.city = round(Num.resumes*100/Total.resumes.this.city,digits=2))

ggplot(num_resumes_per_skill,aes(x=Skill,y=Percent.of.resumes.this.city,fill=City)) + 
geom_bar(stat="identity",position="dodge") + 
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
ylab("Percent of resumes in city with keywords")
```

Based on the plot, it appears that New York has more resumes containing keywords for skills of Collaboration, Communication, and Presentation.

San Francisco resumes seem to contain more keywords for Leadership.

Let's run a chi-squared test for significance for each of these skills.

```{r}
matrix_for_chi_squared_all_skills <- num_resumes_per_skill %>%
				mutate(Num.resumes.without.this.skill = Total.resumes.this.city - Num.resumes) %>%
				select(c("Skill","City","Num.resumes","Num.resumes.without.this.skill"))

colnames(matrix_for_chi_squared_all_skills) <- c("Skill","City","Has","Does.not.have")

for(skill in c("Collaboration","Communication","Presentation","Leadership"))
{
table_for_chi_squared <- matrix_for_chi_squared_all_skills %>% filter(Skill == skill) %>% select(c("Has","Does.not.have"))
rownames(table_for_chi_squared) <- c("New York","San Francisco")
print(skill)
print(table_for_chi_squared)
print(chisq.test(table_for_chi_squared))
}
```

Most of the differences we find are not significant.

However, we do find that keywords associated with skill Collaboration appear in New York data scientist resumes significantly more often than in San Francisco resumes.
