# Most frequent skills in Indeed.com resumes from New York and San Francisco
## Heather Geiger and Raj Kumar - March 25, 2018

### Load libraries and processed data.

```{r, echo=TRUE,eval=TRUE}
library(stringr)
suppressWarnings(suppressMessages(library(tidyverse)))
library(dplyr)
library(ggplot2)
```

```{r, echo=TRUE, eval=TRUE}
load("resumes_across_cities_one_text_field_per_resume.Rdata")
```

### Read in list of skills with keywords and get regular expressions for each.

For each resume, we want to check for the occurence of various strings associated with each skill.

Read in a table with the skill heading plus synonyms.

All in lowercase.

We should require these be bounded on either side by either the start/end of the string, or whitespace, or punctuation.

Start by getting a list of keywords, with a vector of keywords per skill.

```{r, echo=TRUE,eval=TRUE}
keywords <- read.table("keywords.txt",header=TRUE,check.names=FALSE,stringsAsFactors=FALSE,sep="\t")

#In keywords file, we included a few skills we thought would be too tough to look for using a simple word search without doing NLP (natural language processing). For now, since we are not doing NLP here, remove these skills.

keywords <- keywords[grep('This is probably too tough',keywords$Other.notes,invert=TRUE),]

keyword_list <- vector("list",length=nrow(keywords))

for(i in 1:nrow(keywords))
{
keywords_this_row <- keywords$Skill[i]
if(keywords$Synonyms[i] != "None"){
    keywords_this_row <- c(keywords_this_row,unlist(strsplit(keywords$Synonyms[i],",")[[1]]))
    }
keyword_list[[i]] <- keywords_this_row
}
```

Write a function to give a pattern for a keyword if it has a word boundary, comma, or space on each side.

Then if there are multiple keywords, paste these together with a pipe.

Finally, run this function for every item in keyword_list.

```{r, echo=TRUE, eval=TRUE}
#Couldn't figure out how to get a regex for a space, comma, or word boundary.
#However did get one that can get either a space or comma.

space_or_comma <- "[[:space:],]"
word_boundary <- "\\b"

pattern_for_one_keyword <- function(keyword){
    regexes <- paste0(space_or_comma,keyword,space_or_comma)
    regexes <- c(regexes,paste0(word_boundary,keyword,word_boundary))
    regexes <- c(regexes,paste0(word_boundary,keyword,space_or_comma))
    regexes <- c(regexes,paste0(space_or_comma,keyword,word_boundary))
    return(paste0(regexes,collapse="|"))
}

pattern_for_multiple_keywords <- function(keyword_vector){
    if(length(keyword_vector) == 1){return(pattern_for_one_keyword(keyword_vector))}
    if(length(keyword_vector) > 1){
        individual_regexes <- c()
        for(i in 1:length(keyword_vector))
        {
            individual_regexes <- c(individual_regexes,pattern_for_one_keyword(keyword_vector[i]))
        }
    return(paste0(individual_regexes,collapse="|")) 
    }
}

keyword_regexes <- unlist(lapply(keyword_list,function(x)pattern_for_multiple_keywords(x)))
```

### Extracting skills from resumes

We can now use keyword_regexes along with str_detect to give a TRUE/FALSE value for whether the text contains the pattern.

```{r, echo=TRUE, eval=TRUE}
for(i in 1:length(keyword_regexes))
{
skill <- keyword_regexes[i]
skill_in_text <- str_detect(tolower(resumes_across_cities$Text),skill)
if(length(which(skill_in_text == TRUE)) >= 30) #Require that a skill be found in at least 30 resumes including both cities.
{
resumes_across_cities_incl_this_skill <- resumes_across_cities[skill_in_text,]
resumes_across_cities_incl_this_skill <- resumes_across_cities_incl_this_skill %>%
					group_by(City) %>%
					summarize(Num.resumes = n())
resumes_across_cities_incl_this_skill <- data.frame(Skill = keywords$Skill[i],
					data.frame(resumes_across_cities_incl_this_skill),
					stringsAsFactors=FALSE)
if(!(exists("num_resumes_per_skill"))){num_resumes_per_skill <- resumes_across_cities_incl_this_skill;next}
if(exists("num_resumes_per_skill")){num_resumes_per_skill <- rbind(num_resumes_per_skill,resumes_across_cities_incl_this_skill)}
}}
```

Add a column Percent.resumes which expresses the percentage of resumes from the city that contain the skill.

```{r, echo=TRUE, eval=TRUE}
resumes_per_city <- resumes_across_cities %>%
			group_by(City) %>%
			summarize(Total.resumes.this.city = n())
resumes_per_city <- data.frame(resumes_per_city)
resumes_per_city[,1] <- as.vector(resumes_per_city[,1])

num_resumes_per_skill <- merge(num_resumes_per_skill,resumes_per_city,"City") %>% 
			mutate(Percent.resumes = round(Num.resumes*100/Total.resumes.this.city,digits=2))
```

Add information about whether each skill is a soft or technical skill.

```{r, echo=TRUE, eval=TRUE}
num_resumes_per_skill <- merge(num_resumes_per_skill,keywords[,c("Skill","Soft.or.technical")],by="Skill")
```

### Plotting frequency of skills in resumes across cities

#### Soft skills

```{r, echo=TRUE, eval=TRUE}
ggplot(num_resumes_per_skill %>% filter(Soft.or.technical == "soft"),
aes(x=Skill,y=Percent.resumes,fill=City)) +
geom_bar(stat="identity",position="dodge") + 
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
ylab("Percent of resumes with this skill") +
ggtitle("Soft skills in resumes across cities")
```

We find that collaboration, communication, leadership, and presentation are among the top soft skills referenced in data science job seeker resumes.

It also seems like collaboration is mentioned more frequently in New York resumes. 

Let's check if the difference is statistically significant.

```{r, echo=TRUE, eval=TRUE}
percent_resumes_mentioning_collaboration <- num_resumes_per_skill %>% filter(Skill == "collaboration")
percent_resumes_mentioning_collaboration
percent_resumes_mentioning_collaboration <- percent_resumes_mentioning_collaboration %>%
					mutate(Num.resumes.that.do.not.mention = Total.resumes.this.city - Num.resumes) %>%
					select(c("Num.resumes","Num.resumes.that.do.not.mention"))
chisq.test(percent_resumes_mentioning_collaboration)
```

We find that New York data scientist job seekers on Indeed.com use keywords associated with collaboration in their resumes at a statistically significantly higher rate than San Francisco data scientist job seekers.

28.7% of resumes in New York refer to collaboration, compared to only 21.4% of resumes in San Francisco.

#### Technical skills - frequency by city

For technical skills, there are too many skills to have two bars for each skill for all skills.

Let's take the top 10 skills by absolute difference between cities, then plot just these to start.

```{r, echo=TRUE, eval=TRUE}
technical_skills_spread <- num_resumes_per_skill %>% 
			filter(Soft.or.technical == "technical") %>%
			select("Skill","City","Percent.resumes") %>%
			spread("City","Percent.resumes")

colnames(technical_skills_spread)[2:3] <- c("New.York","San.Francisco")

technical_skills_spread <- technical_skills_spread %>%
		mutate(City.difference = New.York - San.Francisco) %>%
		arrange(desc(abs(City.difference)))

technical_skills_spread_top_10 <- technical_skills_spread[1:10,]

colnames(technical_skills_spread_top_10)[2:3] <- c("New York","San Francisco")

technical_skills_top_10_most_different_by_city <- gather(technical_skills_spread_top_10[,1:3],City,Percent.resumes,-Skill)

technical_skills_top_10_most_different_by_city$Skill <- factor(technical_skills_top_10_most_different_by_city$Skill,
							levels=unique(as.vector(technical_skills_top_10_most_different_by_city$Skill)))

ggplot(technical_skills_top_10_most_different_by_city,
aes(x=Skill,y=Percent.resumes,fill=City)) +
geom_bar(stat="identity",position="dodge") +
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
ylab("Percent of resumes with this skill") +
ggtitle("Technical skills in resumes across cities\nTop 10 technical skills with most different frequency between cities")
```

It looks like New York resumes mention data visualization more often than San Francisco resumes.

Again, let's check if the difference is statistically significant.

```{r, echo=TRUE, eval=TRUE}
percent_resumes_mentioning_data_visualization <- num_resumes_per_skill %>% filter(Skill == "data visualization")
percent_resumes_mentioning_data_visualization
percent_resumes_mentioning_data_visualization <- percent_resumes_mentioning_data_visualization %>%
					mutate(Num.resumes.that.do.not.mention = Total.resumes.this.city - Num.resumes) %>%
					select(c("Num.resumes","Num.resumes.that.do.not.mention"))
chisq.test(percent_resumes_mentioning_data_visualization)
```

We do find a statistically significant difference, with data visualization mentioned more frequently in New York resumes than San Francisco resumes (37% vs. 28% of resumes).

#### Technical skills - Python vs. R

One of the questions we were curious about was whether or not Python or R were more popular data science skills.

Let's check the numbers.

```{r, echo=TRUE,eval=TRUE}
num_resumes_per_skill_spread <- num_resumes_per_skill %>%
				select(c("Skill","City","Percent.resumes")) %>%
				spread(City,Percent.resumes)
colnames(num_resumes_per_skill_spread)[2:3] <- c("New.York","San.Francisco")
num_resumes_per_skill_spread <- num_resumes_per_skill_spread %>%
				mutate(Mean.percent.resumes = (New.York + San.Francisco)/2)
num_resumes_per_skill_spread %>% filter(Skill == "python" | Skill == "r")
```

We find that data scientist job seekers more frequently list Python as a skill on their resumes compared to R (64.2% of resumes in each city on average, vs. 55.5% for R).

It will be interesting to see how this compares to job postings.

Does more data scientists knowing Python reflect this skill being more in demand from employers?

Or could less job seekers being fluent in R potentially increase the relative value of R in the job market?

Let's also check the co-occurence of these two skills.

```{r}
resumes_with_python <- str_detect(tolower(resumes_across_cities$Text),keyword_regexes[keywords$Skill == "python"])
resumes_with_R <- str_detect(tolower(resumes_across_cities$Text),keyword_regexes[keywords$Skill == "r"])

round(table(resumes_with_python,resumes_with_R)/nrow(resumes_across_cities)*100,digits=1)
```

We find that a pretty large proportion of all resumes (43.8% if we do not separate by city) include both R and Python. 

The relative probability that a given job seeker will know the other language if they know one of R/Python is thus quite high. 

However, it is more likely for a given job seeker to know Python but not R, than to know R but not Python.

#### Technical skills - frequency across cities

Finally, let's plot the mean frequency for both cities for all technical skills.

```{r, echo=TRUE, eval=TRUE}
technical_skills_to_plot <- technical_skills_spread %>% 
			mutate(Mean.percent.of.resumes = (New.York + San.Francisco)/2)

technical_skills_to_plot <- technical_skills_to_plot %>%
			arrange(Mean.percent.of.resumes)

technical_skills_to_plot$Skill <- factor(technical_skills_to_plot$Skill,
			levels=as.vector(technical_skills_to_plot$Skill))

mycol <- c("#004949","#009292","#FF6DB6","#FFB677","#490092","#006DDB","#B66DFF","#6DB6FF","#B6DBFF","#920000","#924900","#DBD100","#24FF24","#FFFF6D","#000000") #Set up colorblind friendly vector. 

ggplot(technical_skills_to_plot,
aes(x=Skill,y=Mean.percent.of.resumes,fill=Skill)) +
geom_bar(stat="identity") +
ggtitle("Technical skills in resumes across cities") +
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
scale_fill_manual(values=rep(mycol,times=2)) +
ylab("Mean percent of resumes per city")
```

The top 5 technical skills listed by data science job seekers are Python, modeling, SQL, statistics, and R.

Machine learning, prediction, data visualization, data mining, and Java round out the top 10.

Notably, Perl is listed by very few job seekers.
